---
title: "Beaufort Sea chart for Climate Chart Gallery"
author: "Althea Archer"
date: "2024-05-22"
output: html_document
---



## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Define libraries here.

```{r libraries, warning=FALSE, message=FALSE}
# Load libraries
library(tidyverse) # includes ggplot
library(readr) # best for reading and writing csvs
library(sbtools)
library(readxl)

# These are used for the layout of the viz
library(cowplot) # for laying out the final plot
library(sysfonts) # for text editing
library(showtext) # for adding in google fonts
library(magick) # for adding logo
```

## Load files

Sciencebase:

> Gemery, L., 2021, Data Release to Multi-proxy record of ocean-climate variability during the last 2 millennia on the Mackenzie Shelf, Beaufort Sea (2013): U.S. Geological Survey data release, https://doi.org/10.5066/P9SRRW6T.

Original manuscript: 

> Gemery, L., Cronin, T.M., Cooper, L.W., Roberts, L.R., Keigwin, L.D., Addison, J.A., Leng, M.J., Lin, P., Magen, C., Marot, M.E., and Schwartz, V., 2023, Multi-proxy record of ocean-climate variability during the last two millennia on the Mackenzie Shelf, Beaufort Sea: Micropaleontology, v. 69, no. 3, p. 345–360, https://doi.org/10.47894/mpal.69.3.04.

```{r load}
sbtools::item_file_download(sb_id = "6197a410d34eb622f692ce0e",
                            names = "Data release HLY1302 IP135359.xlsx",
                            destinations = "in/Data release HLY1302 IP135359.xlsx",
                            overwrite_file = FALSE)

raw_pct_abundance <- read_excel(path = "in/Data release HLY1302 IP135359.xlsx", 
                                sheet = "T2 Ost MC29bin5G30bin10J32bin20",
                                skip = 1, 
                                range = cell_cols(c(7, 70:131)))

forams_raw <- read_excel(path = "in/HLY1302 faunal bin foraminifera.xlsx", 
                                sheet = "HLY1302 FORAM counts as valu",
                                range = cell_cols(c(7, 70:131))) 

```

## Get data ready for plotting

This next section is for doing any pre-processing steps, data joins, etc, to get your data ready for plotting with ggplot. Remember that most of the time, ggplot prefers data in a "wide" format.

If your processing is expensive or involved, feel free to put in into a separate script or pipeline and read in the processed output here.

```{r processing}
# Clean up data (drop unneeded columns, rename columns)
clean_data <- raw_pct_abundance |> 
  select("Calendar AGE",
         "Kotoracythere arctoborealis Schornikov and Zenina 2006...90",
         "Paracyprideis pseudopunctillata Swain 1963...93") |>
  rename(K_arctoborealis = "Kotoracythere arctoborealis Schornikov and Zenina 2006...90",
         P_pseudopunctillata = "Paracyprideis pseudopunctillata Swain 1963...93",
         YEAR = "Calendar AGE") |>
  mutate(other_species = 100 - (K_arctoborealis + P_pseudopunctillata),
         total_pct = K_arctoborealis + P_pseudopunctillata + other_species,
         YEAR = round(YEAR, 0))

long_pct_data <- clean_data |>
  pivot_longer(cols = c(K_arctoborealis, P_pseudopunctillata, other_species),
               names_to = "species")

# clean up forams
foram_clean_data <- forams_raw |> 
  select("calendar yr",
         "Spiroplectammina biformis...34",
         "Spiroplectammina earlandi...35",
         "E. excavatum clavatum...28") |>
  rename(S_biformis = "Spiroplectammina biformis...34",
         S_earlandi = "Spiroplectammina earlandi...35",
         E_excavatum = "E. excavatum clavatum...28",
         YEAR = "calendar yr") |>
  mutate(
    S_spp = S_biformis + S_earlandi,
    other_species = 100 - (S_spp + E_excavatum),
    total_pct = S_spp + E_excavatum + other_species,
    YEAR = round(YEAR, 0)) |>
  drop_na(YEAR)

foram_long_pct_data <- foram_clean_data |>
  pivot_longer(cols = c(S_spp, E_excavatum, other_species),
               names_to = "species")
```

## Set up main plot

This chunk is where the main ggplot grob definition and set-up occurs. You may have several ggplot grobs, depending on your approach. For each one, if you define it with the `<-` assignment operator and surround the whole code statement in parentheses, you'll be able to preview here what the plot looks like before the next cowplot step.

> Note: The hardest part of this process is getting text to be the optimum size in the final `png`. The font sizes here will **not** be the same as in the final image after the cowplot step. Make sure to check the output `png` for true font sizing and plot composition.

```{r plotting}
# Load some custom fonts and set some custom settings
title_font <- "Lora"
sysfonts::font_add_google(title_font)
supporting_font <- "Source Sans Pro"
sysfonts::font_add_google(supporting_font)
showtext::showtext_opts(dpi = 300, regular.wt = 200, bold.wt = 700)
showtext::showtext_auto(enable = TRUE)
title_font_size <- 55
explainer_font_size <- 18

# Define colors
background_color <- "#ffffff"
font_color <- "#0A1927"
  
# Define margin for cowplotting
margin_cowplot <- 0.04

# The background canvas for your viz on mobile
mobile_height <- 16
mobile_width <- 9
canvas_mobile <- grid::rectGrob(
  x = 0, y = 0, 
  width = mobile_width, height = mobile_height,
  gp = grid::gpar(fill = background_color, alpha = 1, col = background_color)
)

# The background canvas for your viz on instagram
square_dim <- 9
canvas_square <- grid::rectGrob(
  x = 0, y = 0, 
  width = square_dim, height = square_dim,
  gp = grid::gpar(fill = background_color, alpha = 1, col = background_color)
)

# Load in USGS logo (black and white logos available)
usgs_logo <- magick::image_read("../usgs_logo_black.png")
# usgs_logo <- magick::image_read("../usgs_logo_white.png") 

# Main plot
(main_plot <- ggplot(long_pct_data,
                     aes(fill = species,
                         y = value,
                         x = YEAR)) +
    annotate(geom = "rect", 
             xmin = 800, xmax = 1200,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1300, xmax = 1600,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1900, xmax = 2000,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    geom_bar(position = "stack", stat = "identity",
             width = 15, color = NA) +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "top",
          legend.title = element_blank()) +
    xlim(c(2000,0)) +
    scale_fill_manual(values = c("#Ecb05f", "#FDEFDF", "#3C475A"))
 )
ggsave("out/long_time.png",
       height = 20, width = 4)

(foram_main_plot <- ggplot(foram_long_pct_data,
                     aes(fill = fct_rev(species),
                         y = value,
                         x = YEAR)) +
    annotate(geom = "rect", 
             xmin = 800, xmax = 1200,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1300, xmax = 1600,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1900, xmax = 2000,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    geom_bar(position = "stack", stat = "identity",
             width = 15, color = NA) +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "top",
          legend.title = element_blank()) +
    xlim(c(2000,0)) +
    ylim(c(0,100)) +
    scale_fill_manual(values = c("#Ecb05f", "#FDEFDF", "#3C475A"),
                      breaks = c("S_spp", "other_species", "E_excavatum"))
 )
ggsave("out/foram_long_time.png",
       height = 20, width = 4)

# (k_arcto <- ggplot(data = clean_data,
#                    aes(x = YEAR,
#                        y = K_arctoborealis)) +
#     geom_point() +
#     geom_smooth(method = "lm"))
# 
# (P_pseudo <- ggplot(data = clean_data,
#                     aes(x = YEAR,
#                         y = P_pseudopunctillata)) +
#     geom_point() +
#     geom_smooth(method = "lm"))
# 
# # Main plot: forams
# (foram_main_plot <- ggplot(foram_long_pct_data,
#                      aes(fill = species,
#                          y = value,
#                          x = (YEAR))) +
#     geom_bar(position = "stack", stat = "identity") +
#     coord_flip()
#  )
# 
# (s_spp <- ggplot(data = foram_clean_data,
#                    aes(x = YEAR,
#                        y = S_spp)) +
#     geom_point() +
#     geom_smooth(method = "lm"))
# 
# (E_excavatum <- ggplot(data = foram_clean_data,
#                     aes(x = YEAR,
#                         y = E_excavatum)) +
#     geom_point() +
#     geom_smooth(method = "lm"))

(S_spp <- ggplot(data = foram_clean_data,
                    aes(x = YEAR,
                        y = S_spp)) +
    annotate(geom = "rect", 
             xmin = 800, xmax = 1200,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1300, xmax = 1600,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1900, xmax = 2000,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    geom_point(color = "#Ecb05f") +
    geom_smooth(color = "#Ecb05f", fill = "#fdefdf",
                alpha = 0.5) +
    ggtitle("Agglutinated Species (S. spp)") +
    ylim(c(0,100))) +
  ylab("% Abundance") +
    theme_minimal() +
    theme(legend.position = "top",
          legend.title = element_blank())

ggsave("out/agglutinated_Trend.png",
       height = 4, width = 6)

(K_arcto <- ggplot(data = clean_data,
                    aes(x = YEAR,
                        y = K_arctoborealis)) +
    annotate(geom = "rect", 
             xmin = 800, xmax = 1200,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1300, xmax = 1600,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    annotate(geom = "rect", 
             xmin = 1900, xmax = 2000,
             ymin = 0, ymax = 100,
             fill = "grey80", alpha = 0.37) +
    geom_point(color = "#Ecb05f") +
    geom_smooth(color = "#Ecb05f", fill = "#fdefdf",
                alpha = 0.5, method = "lm") +
    ggtitle("K. arctoborealis") +
    #ylim(c(0,100))) +
  ylab("% Abundance") +
    theme_minimal() +
    theme(legend.position = "top",
          legend.title = element_blank()))

ggsave("out/K-arcto_Trend.png",
       height = 4, width = 6)

```

## Produce final plot - MOBILE

Here, use `cowplot` and `ggsave` to create the final viz. This template includes adding the USGS logo, title, text, etc.

```{r cowplot_mobile, fig.width = 9, fig.height = 16}
ggdraw(ylim = c(0,1), # 0-1 scale makes it easy to place viz items on canvas
       xlim = c(0,1)) +
  # a background
  draw_grob(canvas_mobile,
            x = 0, y = 1,
            height = mobile_height, 
            width = mobile_width,
            hjust = 0, 
            vjust = 1) +
  # the main plot
  draw_plot(main_plot,
            x = margin_cowplot,
            y = 0.1,
            height = 0.8,
            width = 0.92) +
  # explainer text
  draw_label("Explainer text\nin multiple\nlines",
             fontfamily = supporting_font,
             x = 1 - margin_cowplot,   
             y = margin_cowplot,
             size = explainer_font_size,
             hjust = 1,
             vjust = 0,
             color = font_color)+
  # Title
  draw_label("Title",
             x = margin_cowplot,
             y = 1 - margin_cowplot,
             hjust = 0,
             vjust = 1,
             lineheight = 0.75,
             fontfamily = title_font,
             fontface = 'bold',
             color = font_color,
             size = title_font_size) +
  # Add logo
  draw_image(usgs_logo, 
             x = margin_cowplot,
             y = margin_cowplot,
             width = 0.2, 
             hjust = 0, 
             vjust = 0, 
             halign = 0, 
             valign = 0)

# Save the final image
ggsave(filename = "out/yourproject_yourname_mobile.png", 
       width = mobile_width, height = mobile_height, dpi = 300)
```

## Produce final plot - SQUARE

Here, use `cowplot` and `ggsave` to create the final viz. This template includes adding the USGS logo, title, text, etc.

```{r cowplot_square, fig.width = 9, fig.height = 9}
ggdraw(ylim = c(0,1), # 0-1 scale makes it easy to place viz items on canvas
       xlim = c(0,1)) +
  # a background
  draw_grob(canvas_square,
            x = 0, y = 1,
            height = square_dim, width = square_dim,
            hjust = 0, vjust = 1) +
  # the main plot
  draw_plot(main_plot,
            x = margin_cowplot,
            y = 0.11,
            height = 0.78,
            width = 0.92) +
  # explainer text
  draw_label("Explainer text\nin multiple lines",
             fontfamily = supporting_font,
             x = 1 - margin_cowplot,   
             y = margin_cowplot,
             size = explainer_font_size,
             hjust = 1,
             vjust = 0,
             color = font_color)+
  # Title
  draw_label("Title",
             x = margin_cowplot,
             y = 1 - margin_cowplot,
             hjust = 0,
             vjust = 1,
             lineheight = 0.75,
             fontfamily = title_font,
             fontface = 'bold',
             color = font_color,
             size = title_font_size) +
  # Add logo
  draw_image(usgs_logo, 
             x = margin_cowplot,
             y = margin_cowplot,
             width = 0.15, 
             hjust = 0, vjust = 0, 
             halign = 0, valign = 0)

# Save the final image
ggsave(filename = "out/yourproject_yourname_square.png", 
       width = square_dim, height = square_dim, dpi = 300)
```

## Supporting information

### Caption with key message relating this viz

1.  Key takeaways here (1-2 sentences each)

### Data source(s)

Data sources here

### Alt text

Alt text for visualization here

### Authors

List authors here
