---
title: "{Project name} chart for Climate Chart Gallery"
author: "Hayley Corson-Dosch"
date: "2024-04-30"
output: html_document
---

## Instructions for using this template

> Make sure you have copied the template subdirectory and named it with your project name and your name (e.g., `/FishAsFood_hcorson-dosch`) before editing this document!

1.  Put all comments either in the html text (like this) or in the `r` chunks (delineated with 3 "\`" marks) to explain your process to Vizlab reviewers
2.  Put all R code in the chunks. They're organized for you to (in order): do some set-up steps (load libraries), load in any data (manually or with built in packages), process your data for plotting, create plot using ggplot, and finally compose the final image with cowplot.
3.  You can run individual code chunks while developing code and plots. Note: the plots that are created in the RStudio coding window *do not* match the plot as created through the ggsave function.
4.  To preview the final plot, use the `Knit` button above. That will create a preview of the final plot in the output `html` and will save the `png` in the `out/` folder. On many systems, the preview in the script document and the html will not match the output `png`, so please refer to the `png` for final proofing.
5.  When you're happy with the final plot, fill out the supporting information at the bottom and push to gitlab for review. **Note: only commit input data if it can't be downloaded directly with code (e.g., `sbtools` or `dataRetrieval`)**.

This website has great information on how to use RMarkdown: <https://yihui.org/knitr/>

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Define libraries here.

```{r libraries, warning=FALSE, message=FALSE}
# Load libraries
library(tidyverse) # includes ggplot
library(readr) # best for reading and writing csvs

# These are used for the layout of the viz
library(cowplot) # for laying out the final plot
library(sysfonts) # for text editing
library(showtext) # for adding in google fonts
library(magick) # for adding logo
```

## Load files

Save any files that you are using directly in the `in` folder. Then read them in to the environment with this chunk of code. Remember, use the RStudio project and relative file pathways, and never include any personal computer file pathways.

This is also a good place to load any data from other built-in packages such as `dataRetrieval` or `spData`, etc.

```{r load}
# Add your code here



```

## Get data ready for plotting

This next section is for doing any pre-processing steps, data joins, etc, to get your data ready for plotting with ggplot. Remember that most of the time, ggplot prefers data in a "wide" format.

If your processing is expensive or involved, feel free to put in into a separate script or pipeline and read in the processed output here.

```{r processing}
# Add your processing code here


```

## Set up main plot

This chunk is where the main ggplot grob definition and set-up occurs. You may have several ggplot grobs, depending on your approach. For each one, if you define it with the `<-` assignment operator and surround the whole code statement in parentheses, you'll be able to preview here what the plot looks like before the next cowplot step.

> Note: The hardest part of this process is getting text to be the optimum size in the final `png`. The font sizes here will **not** be the same as in the final image after the cowplot step. Make sure to check the output `png` for true font sizing and plot composition.

```{r plotting}
# Load some custom fonts and set some custom settings
title_font <- "Lora"
sysfonts::font_add_google(title_font)
supporting_font <- "Source Sans Pro"
sysfonts::font_add_google(supporting_font)
showtext::showtext_opts(dpi = 300, regular.wt = 200, bold.wt = 700)
showtext::showtext_auto(enable = TRUE)
title_font_size <- 55
explainer_font_size <- 18

# Define colors
background_color <- "#ffffff"
font_color <- "#0A1927"
  
# Define margin for cowplotting
margin_cowplot <- 0.04

# The background canvas for your viz on mobile
mobile_height <- 16
mobile_width <- 9
canvas_mobile <- grid::rectGrob(
  x = 0, y = 0, 
  width = mobile_width, height = mobile_height,
  gp = grid::gpar(fill = background_color, alpha = 1, col = background_color)
)

# The background canvas for your viz on instagram
square_dim <- 9
canvas_square <- grid::rectGrob(
  x = 0, y = 0, 
  width = square_dim, height = square_dim,
  gp = grid::gpar(fill = background_color, alpha = 1, col = background_color)
)

# Load in USGS logo (black and white logos available)
usgs_logo <- magick::image_read("../usgs_logo_black.png")
# usgs_logo <- magick::image_read("../usgs_logo_white.png") 

# Main plot
(main_plot <- ggplot()
 )



```

## Produce final plot - MOBILE

Here, use `cowplot` and `ggsave` to create the final viz. This template includes adding the USGS logo, title, text, etc.

```{r cowplot_mobile, fig.width = 9, fig.height = 16}
ggdraw(ylim = c(0,1), # 0-1 scale makes it easy to place viz items on canvas
       xlim = c(0,1)) +
  # a background
  draw_grob(canvas_mobile,
            x = 0, y = 1,
            height = mobile_height, 
            width = mobile_width,
            hjust = 0, 
            vjust = 1) +
  # the main plot
  draw_plot(main_plot,
            x = margin_cowplot,
            y = 0.1,
            height = 0.8,
            width = 0.92) +
  # explainer text
  draw_label("Explainer text\nin multiple\nlines",
             fontfamily = supporting_font,
             x = 1 - margin_cowplot,   
             y = margin_cowplot,
             size = explainer_font_size,
             hjust = 1,
             vjust = 0,
             color = font_color)+
  # Title
  draw_label("Title",
             x = margin_cowplot,
             y = 1 - margin_cowplot,
             hjust = 0,
             vjust = 1,
             lineheight = 0.75,
             fontfamily = title_font,
             fontface = 'bold',
             color = font_color,
             size = title_font_size) +
  # Add logo
  draw_image(usgs_logo, 
             x = margin_cowplot,
             y = margin_cowplot,
             width = 0.2, 
             hjust = 0, 
             vjust = 0, 
             halign = 0, 
             valign = 0)

# Save the final image
ggsave(filename = "out/yourproject_yourname_mobile.png", 
       width = mobile_width, height = mobile_height, dpi = 300)
```

## Produce final plot - SQUARE

Here, use `cowplot` and `ggsave` to create the final viz. This template includes adding the USGS logo, title, text, etc.

```{r cowplot_square, fig.width = 9, fig.height = 9}
ggdraw(ylim = c(0,1), # 0-1 scale makes it easy to place viz items on canvas
       xlim = c(0,1)) +
  # a background
  draw_grob(canvas_square,
            x = 0, y = 1,
            height = square_dim, width = square_dim,
            hjust = 0, vjust = 1) +
  # the main plot
  draw_plot(main_plot,
            x = margin_cowplot,
            y = 0.11,
            height = 0.78,
            width = 0.92) +
  # explainer text
  draw_label("Explainer text\nin multiple lines",
             fontfamily = supporting_font,
             x = 1 - margin_cowplot,   
             y = margin_cowplot,
             size = explainer_font_size,
             hjust = 1,
             vjust = 0,
             color = font_color)+
  # Title
  draw_label("Title",
             x = margin_cowplot,
             y = 1 - margin_cowplot,
             hjust = 0,
             vjust = 1,
             lineheight = 0.75,
             fontfamily = title_font,
             fontface = 'bold',
             color = font_color,
             size = title_font_size) +
  # Add logo
  draw_image(usgs_logo, 
             x = margin_cowplot,
             y = margin_cowplot,
             width = 0.15, 
             hjust = 0, vjust = 0, 
             halign = 0, valign = 0)

# Save the final image
ggsave(filename = "out/yourproject_yourname_square.png", 
       width = square_dim, height = square_dim, dpi = 300)
```

## Supporting information

### Caption with key message relating this viz

1.  Key takeaways here (1-2 sentences each)

### Data source(s)

Data sources here

### Alt text

Alt text for visualization here

### Authors

List authors here
